---
title: "Covid Data Analysis"
output: html_document
author: Sam Cohen
---

## Setup
```{r setup, include=TRUE, message = FALSE, warning = FALSE}
# Load packages used in this session of R
library(tidyverse)
library(readxl)
library(openxlsx)
library(dplyr)
library(readr)

# Setting / checking working directory
getwd()

# Reading in data
instability <- load("Ch7_Exercise1_Instability.RData") ## Instability
instability <- dta
ox_covid <- read_csv("OxCGRT_latest.csv") ## Covid Data
ptreat <- readLines("Ch7_Exercise4_PostTreatmentSimulation.R") ## Post treatment simulation code
us_states <- read_excel("USstates.xlsx")  ## US states


```


## Exploratory Analysis

#### Data wrangling

```{r}
# Viewing Covid data
head(ox_covid)
```

```{r}
# cleaning Covid data
ox_covid_us <- ox_covid %>% 
  mutate("stAbbrev" = str_replace_all(string = RegionCode, pattern = "US_",
                                      replacement = "" )) %>%  
  filter(CountryName == "United States") %>% # Filtering for US
  select(RegionName, RegionCode, Date, GovernmentResponseIndex, ConfirmedCases,
         ConfirmedDeaths) # Selecting variables 
  
ox_covid_us <- data.frame(ox_covid_us) # Saving back to data frame
```

```{r}
# Summarizing by state
ox_covid_state <- ox_covid_us %>% 
  dplyr::filter(Date >20201101 & Date < 20201115) %>%  # Filtering by start/end date
  select(RegionName, GovernmentResponseIndex, ConfirmedDeaths, ConfirmedCases) %>% 
  #selecting variables
  group_by(state= RegionName) %>% #grouping by state
  summarize(Average_state_policy = mean(GovernmentResponseIndex), total_deaths = sum(ConfirmedDeaths), total_cases = sum(ConfirmedCases))

#summarizing by Covid
ox_covid_state <-data.frame(ox_covid_state)
```

```{r}
# Merging data sets 

covid_master <- left_join(ox_covid_state, us_states, by="state") %>%
  mutate(death_per_10k = (total_deaths/pop2019)*10000, pop_density = pop2019/sqMiles)
covid_master <- data.frame(covid_master)

#first three variables and first three lines
covid_master[,1:3] %>% top_n(3)
```

#### Plotting deaths per 10,000 people against density.

```{r}
# plotting deaths/10k and density
ggplot(data=covid_master, mapping = aes(x=pop_density, y=death_per_10k
)) +
  geom_point(color="red") +
  labs(x = "Population Density", y = "Deaths per 10k People") +
  geom_smooth(method=lm, se=FALSE) 

```


#### Plotting cases per 10,000 people against density with a color indicating political party of governor. 

```{r }
ggplot(data=covid_master, mapping = aes(x=pop_density, y=death_per_10k
, color = Party)) +
  geom_point() +
  labs(x = "Population Density", y = "Deaths per 10k People") +
  geom_smooth(method=lm, se=FALSE)
```

#### Positive cases as a function of policy, colored by party of governor.

```{r }
ggplot(data=covid_master, mapping = aes(x=Average_state_policy, y=total_cases
, color = Party)) +
  geom_point() +
  labs(x = "Average state policy", y = "COVID Cases") +
  geom_smooth(method=lm, se=FALSE) +
  scale_y_continuous(limits = c(0, 4000000))
```
#### Column plot of deaths per 10k people by State and median age
```{r}
# Reordering states by death/10k count
covid_master$state <- reorder(covid_master$state, -covid_master$death_per_10k)

# Generating column plot
ggplot(data=covid_master, mapping = aes(x=state, y=death_per_10k
, fill = MedianAge)) + # shading by median age of State
  geom_col() +
  labs(x = "State", y = "Deaths per 10k People") + 
  theme(axis.text.x = element_text(angle = 70, hjust = 1)) # Rotating x-axis text


```

#### Scatterplot: Average state policy and deaths per 10k
```{r}
ggplot(data=covid_master, mapping = aes(x=Average_state_policy, y=death_per_10k)) +
         geom_point() +
         labs(x = "Average state policy", y = "Deaths per 10k People") +
         geom_smooth(method=lm, se=FALSE)
  
```

-

## Hypothesis testing

### Regression: relationship between deaths and state policy (with control variables)
```{r}
# Bivariate
lm(death_per_10k ~ Average_state_policy, data = covid_master) %>% summary()

# Multivariate (controlling for age, cases and population density)
lm(death_per_10k ~ Average_state_policy + MedianAge + total_cases + pop_density, data = covid_master) %>% summary()

```

### Estimating a bivariate model with instability as the dependent variable and democracy as the independent variable. Standardized coefficients used. 

```{r}
# Model: democracy and instability
lm(scale(instab) ~ scale(democracy), data=instability)
summary(lm(scale(instab) ~ scale(democracy), data=instability))
```

In the model above, instability is dependent on democracy (standardized). The results show a negative relationship-- the more democratic countries are, the less unstable they are. Specifically, a one positive unit change in democracy leads to a -4.541e-02 difference in standard deviations from the mean. The results are statistically significant.

However, there is likely endogeneity, since there are may factors that can affect instability, such as economic conditions, health, natural disasters, war, etc. 


### Including variable (lagged GDP) to combat endogeneity

```{r}
# Model: democracy + GDPlag and instability

## Turning ln(gdp) to gdp
instability$gdplag <- exp(instability$lngdplag)

lm(scale(instab) ~ scale(democracy) + scale(gdplag), data=instability)
summary(lm(scale(instab) ~ scale(democracy) + scale(gdplag), data=instability)
)
```
Including standardized GDP lag changes the model-- democracy (standardized) now has a positive relationship with instability. However, it still gives a statistically significant result.

### Estimating a model with logged-lagged GDP

```{r}
# Model: democracy + lnGDPlag and instability
lm(scale(instab) ~ scale(democracy) + scale(lngdplag), data=instability)
summary(lm(scale(instab) ~ scale(democracy) + scale(lngdplag), data=instability)
)

lm(scale(instab) ~ scale(democracy) + scale(lngdplag), data=instability)$coefficients
```
